### 2025-02-12 chapters/analysis_techniques.tex
Changes 1. Initialized repo 2. Added current progress with dissertation. Intro, methods, analysis techniques, literature review and some results present

\section{Analysis techniques}

In this work, the structure factor derived domain size, geometric pore size distribution and tortuosity will be utilized to characterize the microstructure. The nematic order parameter will be used to characterize the particle order to the magnetic field. The average angle to the interface will be utilized to characterize how the particles ordering to the field affect the interface position. Finally, the radial distribution function will be used to probe how differences in ordering change the packing of particles on the interface. The time evolution of viscosity will provide insight into the rheological properties of the bijel, while the aforementioned analysis tools in addition to proportion of particles on the interface will be calculated to provide insight into how the bijel reacts to shear, parallel and perpendicular to the magnetic field. A more detailed description of the implementation of each method is shown in subsequent sections.

\subsection{Particle filling and interface smoothening}
\label{section:filling_routine}

Particles are the cause of bijels but they can also make characterization of domain size, curvature, pore size distribution and tortuosity more difficult as particles may change these parameters. An extrapolation scheme is required to connect the interface one different sides of the particles and to fill in the particles themselves so as to create a more accurate interface for other techniques.

This is accomplished through sequential density averaging and filling of grid points making up the particles from the outside in. The grid cells around a cell of interest are averaged using a D3Q19 like lattice and stepping through all points with a order parameter $\phi = \rho^w - \rho^o$ value of 0.

\subsection{Domain size}

The characteristic length scale at time $t$ of a bijel can be computed from the moments of the structure factor. The structure factor is first calculated from the order parameter of the phases, defined as $\phi = \rho_b - \rho_r$ where $\phi$ is the density order parameter and $\rho_b$ and $\rho_r$ are the densities of the blue and red fluid respectively. Next, the fluctuations of the order parameter $\phi'$ is calculated by subtracting the mean of the order parameter, $\langle \phi \rangle$ from the order parameter, defined as $\phi' = \phi - \langle \phi \rangle$. Next, the fourier transform of the fluctuations of the order parameter $\phi'(\mathbf{k})$ is calculated. Finally, $\phi'(\mathbf{k})$ is squared and made positive and normalized by the volume of the system defined as $S(\mathbf{k}) = \frac{1}{N} |\phi'(\mathbf{k})^2|$.

The first domain size metric used is the domain size derived from the spherically averaged first moment of the structure factor. The spherical averaging is done by averaging the values of the structure factor at the same $\mathbf{k}$ up till the nyquist frequency. The spherically averaged structure factor is defined as $S(k)$ where $k$ represents the wave numbers obtained from the spherical averaging. Next, the first moment $\langle k_1 \rangle$ is calculated as $\langle k_1 \rangle = \frac{\Sigma_{k} k\cdot S(k)}{\Sigma_{k} S(k)}$. Finally, the domain size is calculated as $L_1 = \frac{2 \pi}{\langle k_1 \rangle}$. This domain size is used in past literature \textcolor{blue}{https://doi.org/10.1017/S0022112001004682, https://www.doi.org/10.1103/PhysRevE.67.046304} to calculate the correct scaling regime of a system undergoing spinodal decomposition.

While it cannot be used to correlate physical phenomena the second moment of the structure factor derived domain size is utilized here to discriminate directional length scales. Owing to the spherical averaging done, the first moment moment derived domain size is unable to be used to calculate direction specific length scales. The second moment of the structure factor is first calculated by multiplying $\langle \mathbf{k_2}^2 \rangle = \frac{\Sigma_{\mathbf{k}} \mathbf{k}^2 S(\mathbf{k})}{\Sigma_k S(\mathbf{k})}$ where $\mathbf{k_2}^2$ represents the second moment in each cartesian direction. Next, the domain size in each direction is calculated as $L_2 = \frac{2 \pi}{\sqrt{\mathbf{k_2}^2}}$. \cite{jansen_bijels_2011, gunther_timescales_2014} In the rest of this document, $L_{\parallel}$ and $L_{\perp}$ is defined as the domain size parallel and perpendicular to the applied magnetic field.

\subsection{Tortuosity}

Tortuosity can be defined as the ratio between the path length $(L_{eff})$ of two points and the euclidean distance between two points $L_{0}$, $\tau = L_{eff}/L_{0}$. In practice this is the geometric definition of tortuosity as often the tortuosity of a system is dependent upon other factors. Diffusive tortuosity is what we will use here, defined as $\tau = \frac{D_{eff}}{D_0}$. This metric is also more relevant to the potential applications of bijels which rely on diffusion of mass for their uses.

Simulation techniques such as Lattice Boltzmann or Molecular Dynamics have been used to characterize the tortuosity of porous materials. However, these simulations can be computationally expensive and time consuming to perform, even though they may have great accuracy. Recently, particle diffusion techniques based on a random walk such as pytrax and taufactor or Pore Network Modelling (PNM) based techniques such as PoreSpy have been introduced which offer a tradeoff of accuracy for speed. Taufactor in particular is implemented on GPU's with a periodic boundary condition, allowing for rapid and accurate calculation of the tortuosity.

\textcolor{blue}{https://joss.theoj.org/papers/10.21105/joss.05358, https://doi.org/10.1016/j.softx.2019.100277, https://joss.theoj.org/papers/10.21105/joss.01296}

Before using taufactor, the microstructure is first binarized. This is done by first calculating the order parameter of the system with $\phi = \rho_b - \rho_r$ where $\phi$ is the order parameter and $\rho_b, \rho_r$ are the densities of the blue and red fluid respectively. Next, the interface described with $\phi$ is smoothened using the algorithm detailed in section \ref{section:filling_routine}. In many of the bijels synthesized using small molecule liquids, monomer is introduced into the oil phase and polymerized. Therefore, to perform the binarization step, we select all points where $\phi \leq 0$ and set their value to $0$, and all other values are set to $1$. To taufactor, $1$ represents void spaces and 0's are solid.

\subsection{Nematic order parameter}

The orientational ordering of the particles to one another can be characterized using the nematic order parameter, defined as $S$. This parameter was first used to characterize liquid crystals. However its use has expanded to colloidal systems to characterize the ordering of a system to an average orientation, termed the director. A system is said to have nematic ordering if its components have orientational order to the director but does not require the system to have translational order.

The nematic order parameter is calculated from the orientational alignment tensor, $Q$, which is calculated from the orientation of particles to one another as $\langle Q \rangle = \frac{1}{N} \sum_{i}^{N} \langle \frac{3}{2}u_i u_i - \frac{1}{2}I \rangle$. $u_i$ is the orientation of particle $i$ and $I$ is an identity matrix. \cite{veerman_phase_1992} While there are multiple methods to evaluate the nematic order parameter, the one we have picked is where the largest eigenvector is $S$, with its corresponding eignvector referring to the director of the system. \cite{veerman_phase_1992} \textcolor{blue}{https://doi.org/10.1080/00268978400101951} A system is defined to be ordered nematically, if $S > 0.3$. In this work, the direction of the director will be compared to the direction of the magnetic field. We utilize the implementation in Freud to calculate the nematic order parameter.

\subsection{Radial distribution function}

The radial distribution function (RDF) is used to measure the normalized number density of particles around a reference particle as a function of distance. The normalization factor is the expected density of particles derived from an ideal gas. The general form of the equation to calculate the RDF is $g(r) = \frac{1}{\rho_{ideal}(r) 4\pi r^2} \int_{r}^{r+\Delta r} \delta(r_{i} - r_{j}) dr$ where $\rho_{ideal}(r)$ represents the density expected at distance r from a reference particle in an ideal gas, $\delta(r_{i} - r_{j})$ represents the number of particles within the bin $r + \Delta r$ and integrated to get a volumetric density. As implemented in a numerical method, the integral becomes a sum. After correcting for the periodic boundary condition using the minimum image convention, the RDF can be used to characterize periodic systems. We utilize the implementation in the software package Freud.

\subsection{Average interface angle and proportion of particles on the interface}

One way to differentiate between whether the particles pull the interface or whether the tilt of the particles at the interface causes changes to the jamming point is to characterize the average interface angle. The average interface angle $(\psi)$ is defined as the angle between the orientation of the axis of symmetry of the particle and the interface normal. The former mechanism would show little to no change in the interface angle, while the former mechanism should show large differences. The proportion of particles on the interface can also be calculated using this technique.

The orientation of the particle is obtained from the output files from the simulation. The interface normal needs to be calculated from $\phi$. This is done by taking $\phi$ as calaculated from the red and blue fluid densities, then performing the particle fill routine described in section \ref{section:filling_routine}. Next, a mesh corresponding to the interface is obtained using a marching cubes algorithm at a countour level of 0 with a cube length of 4 lattice units, which corresponds to the thickness of the interface. \textcolor{blue}{https://doi.org/10.7717/peerj.453} The position of the particle and its closest vertex is calculated. If this distance is larger than the first minimum in the radial distribution function, it is considered to be not on the interface and the next step is not performed. The vertex and corresponding normal closest to the position of a particle is used in the dot product, $\cos{\psi} = \frac{n_{int} \cdot o_{i}}{|n_{int}| |o_i|}$ where $n_{int}$ is the normal of the interface and $o_i$ is the orientation of particle i. After looping through all the particles, the average of $\psi$ and the number of particles in the interface is calculated.

\subsection{Shear viscosity}
\label{section:shear_viscosity}

Shear viscosity is a physical parameter measured through the dissipation of applied strain onto a fluid, resulting in an applied shear stress. In an incompressible fluid, the viscosity $(\eta)$ is defined as the ratio between the measured shear stress $(\sigma)$ and the applied strain rate $(\dot{\gamma})$, $\sigma = \eta \dot{\gamma}$. In rheological experiments, the shear viscosity is measured through a viscometer based on Couette or shear viscometry and Poiselle or pressure gradient viscometry. A Couette cell functions by applying a shear velocity to one or more walls, and measuring the shear stress response of the fluid. Common geometries for Couette cells include the cone and plate and concentric cylinder viscometers.

In this work, a shear velocity is imparted in the z direction, across the x-axis. Therefore, the shear stress in the x and z directions, $\sigma_{xz}$, is measured. $\sigma_{xz}$ is output per lattice in the output file, which is then averaged across the entire volume to obtain an average shear stress $\hat{\sigma_{xz}}$. The effective viscosity of the system, $\eta_{eff}$ is calculated using $\eta_{eff} = \frac{\hat{\sigma_{xz}}}{\dot{\gamma}}$ where $\dot{\gamma} = \frac{2 u_{LE}}{L_x}$. $u_{LE}$ is the shear velocity is enforced through the Lees-Edwards boundary condition, as described in section \textcolor{blue}{description of lees edwards BC}.

\subsection{Yield stress and zero shear rate viscosity}

The yield stress of bijels can be calculated using the flow curve, stress ramp or from oscillatory rheological techniques. The flow curve method involves the steady application of shear, followed by measurements of the shear stress response of the material. The plateau of the shear stress response is indicative of the yield stress of the material. The stress ramp method involves the application of a shear stress onto the material, followed by measurements of the strain rate. A perfectly elastic material will have no strain rate, while a perfectly disspitative material will have high strain rate. The yield stress is characterized as the point when a rapid increase in strain rate is observed. The yield stress from oscillatory rheology is defined when the storage modulus is lower than the loss modulus. This work will utilize the flow curve method.

The shear viscosity calculated using the method detailed in Section \ref{section:shear_viscosity} will be plotted as a function of applied shear rate. Rheological models such as the Herschel Buckley, Bingham Plastic or Carreu models can be utilized to fit the shear viscosity as a function of shear rate to calculate the zero shear rate viscosity. In bijel applications, this parameter gives insight into the material stability and processability.