### 2025-02-12 licenses/licenses.tex
Changes 1. Created license folder and added list of licenses currently or to be used 2. Modified gitignore file to ensure that latex make files are not added to git

\section{License numbers}

\begin{itemize}
    \item Continuous Fabrication of Hierarchical and Asymmetric Bijel Microparticles, 
    Fibers, and Membranes by Solvent Transfer‐Induced Phase Separation (STRIPS): 5913140219015
    \item Scalable synthesis of gyroid-inspired freestanding three-dimensional graphene architectures: 1548092-1
    \item State diagram for particle stabilized emulsions. [Soft Matter, 2015,11, 8393-8403]: 1551968-1
    \item Scalable Manufacturing of Hierarchical Biphasic Bicontinuous Structures via Vaporization-Induced 
    Phase Separation (VIPS) [Materials Letters 2020, 2, 5, 524-530]: "Reprinted (adapted) with permission from 
    {COMPLETE REFERENCE CITATION}. Copyright {YEAR} American Chemical Society."
    \item Modulation of Cellulose Nanocrystals Amphiphilic Properties to Stabilize Oil/Water Interface 
    [Biomacromolecules 2016, 17, 5, 1748–1756]: "Reprinted (adapted) with permission from 
    {COMPLETE REFERENCE CITATION}. Copyright {YEAR} American Chemical Society."
    \item Dextran-Based Nanoparticles to Formulate pH-Responsive Pickering Emulsions: A Fully Degradable 
    Vector at a Day Scale [Biomacromolecules 2020, 21, 12, 5358–5368]: "Reprinted (adapted) with permission from 
    {COMPLETE REFERENCE CITATION}. Copyright {YEAR} American Chemical Society."
    \item Stable emulsions with thermally responsive microstructure and rheology using poly(ethylene oxide) star 
    polymers as emulsifiers [Journal of Colloid and Interface Science, 2013, 394, 284-292]: 5921030909190
    \item Magnetophoresis of Magnetic Pickering Emulsions Under Low Field Gradient: Macroscopic and Microscopic 
    Motion [Langmuir 2021, 37, 5, 1811–1822]: "Reprinted (adapted) with permission from 
    {COMPLETE REFERENCE CITATION}. Copyright {YEAR} American Chemical Society."
    \item Influence of magnetic field on the orientation of anisotropic magnetic particles at liquid interfaces 
    [Phys Chem Chem Phys 2014, 16, 47, 26051-26058]: 1551970-1
    \item Pickering Emulsions with Controllable Stability [Langmuir, 2005, 21, 6, 2158-2162]: "Reprinted (adapted) with permission from 
    {COMPLETE REFERENCE CITATION}. Copyright {YEAR} American Chemical Society."
    \item Bijels Containing Magnetic Particles: A Simulation Study [Langmuir, 2010, 26, 11, 7928-7936]: 
    "Reprinted (adapted) with permission from {COMPLETE REFERENCE CITATION}. Copyright {YEAR} 
    American Chemical Society."
    \item Numerical simulations of particulate suspensions via a discretized Boltzmann equation. Part 2. Numerical results
    \item Aerobijels: Ultralight Carbon Monoliths from Cocontinuous Emulsions 
    [Adv. Funct. Mater. 2020, 30, 1908383]: 5921080017718
    \item Hierarchical assemblies of superparamagnetic colloids in 
    time-varying magnetic fields [Soft Matter. 17, 5, 1120-1155]: 1552854-1
\end{itemize}