### 2025-03-10 licenses/licenses.tex

\section{License numbers}

\begin{itemize}
    \item E. Garcia, C. S. Wang, R. N. Sanderson, K. M. McDevitt, Y. Zhang, L. Valdevit, D. R.
    Mumm, A. Mohraz and R. Ragan. Scalable synthesis of gyroid-inspired freestanding three-
    dimensional graphene architectures. 1, 3870.: 1548092-1
    \item Stratford, R. Adhikari, I. Pagonabarraga, J.-C. Desplat and M. E. Cates. Colloidal jam-
    ming at interfaces: A route to fluid-bicontinuous gels. 309, 2198: 5966820525314
    \item F. Haase, K. J. Stebe and D. Lee. Continuous fabrication of hierarchical and asymmetric
    bijel microparticles, fibers, and membranes by solvent transfer-induced phase separation
    (STRIPS). 27, 7065: 5913140219015
    \item K. Tham, W. M. Ng, S. S. Leong, S. P. Yeap, S. C. Low, H. L. Lee and J. Lim. Mag-
    netophoresis of magnetic pickering emulsions under low field gradient: Macroscopic and
    microscopic motion. 37, 1811: "Reprinted (adapted) with permission from 
    Langmuir 2021, 37, 5, 1811-1822. Copyright 2025 American Chemical Society."
    \item G. B. Davies, T. Kr¨uger, P. V. Coveney, J. Harting and F. Bresme. Assembling ellipsoidal
    particles at fluid interfaces using switchable dipolar capillary interactions. 26, 6715.: Reproduced under the Creative Commons CC BY License.
    \item C. Loudet, A. M. Alsayed, J. Zhang and A. G. Yodh. Capillary interactions between
    anisotropic colloidal particles. 94, 018301: RNP/25/FEB/088185
    \item S. S. Velankar. A non-equilibrium state diagram for liquid/fluid/particle mixtures. 11, 8393: 1551968-1
    \item Bonaccorso, S. Succi, M. Lauricella, A. Montessori, A. Tiribocchi and K. H. Luo. Shear
    dynamics of confined bijels. 10, 095304.: Reproduced under the Creative Commons CC BY License.
    \item S. Schmieschek, L. Shamardin, S. Frijters, T. Kr¨uger, U. D. Schiller, J. Harting and P. V.
    Coveney. LB3D: A parallel implementation of the Lattice-Boltzmann method for simulation
    of interacting amphiphilic fluids. Computer Physics Communications 217, 149 (2017).: Reproduced under the Creative Commons CC BY License
\end{itemize}

% \begin{itemize}
%     \item Continuous Fabrication of Hierarchical and Asymmetric Bijel Microparticles, 
%     Fibers, and Membranes by Solvent Transfer‐Induced Phase Separation (STRIPS): 5913140219015
%     \item Scalable synthesis of gyroid-inspired freestanding three-dimensional graphene architectures: 1548092-1
%     \item State diagram for particle stabilized emulsions. [Soft Matter, 2015,11, 8393-8403]: 1551968-1
%     \item Scalable Manufacturing of Hierarchical Biphasic Bicontinuous Structures via Vaporization-Induced 
%     Phase Separation (VIPS) [Materials Letters 2020, 2, 5, 524-530]: "Reprinted (adapted) with permission from 
%     {COMPLETE REFERENCE CITATION}. Copyright {YEAR} American Chemical Society."
%     \item Modulation of Cellulose Nanocrystals Amphiphilic Properties to Stabilize Oil/Water Interface 
%     [Biomacromolecules 2016, 17, 5, 1748–1756]: "Reprinted (adapted) with permission from 
%     {COMPLETE REFERENCE CITATION}. Copyright {YEAR} American Chemical Society."
%     \item Dextran-Based Nanoparticles to Formulate pH-Responsive Pickering Emulsions: A Fully Degradable 
%     Vector at a Day Scale [Biomacromolecules 2020, 21, 12, 5358–5368]: "Reprinted (adapted) with permission from 
%     {COMPLETE REFERENCE CITATION}. Copyright {YEAR} American Chemical Society."
%     \item Stable emulsions with thermally responsive microstructure and rheology using poly(ethylene oxide) star 
%     polymers as emulsifiers [Journal of Colloid and Interface Science, 2013, 394, 284-292]: 5921030909190
%     \item Magnetophoresis of Magnetic Pickering Emulsions Under Low Field Gradient: Macroscopic and Microscopic 
%     Motion [Langmuir 2021, 37, 5, 1811–1822]: "Reprinted (adapted) with permission from 
%     {COMPLETE REFERENCE CITATION}. Copyright {YEAR} American Chemical Society."
%     \item Influence of magnetic field on the orientation of anisotropic magnetic particles at liquid interfaces 
%     [Phys Chem Chem Phys 2014, 16, 47, 26051-26058]: 1551970-1
%     \item Pickering Emulsions with Controllable Stability [Langmuir, 2005, 21, 6, 2158-2162]: "Reprinted (adapted) with permission from 
%     {COMPLETE REFERENCE CITATION}. Copyright {YEAR} American Chemical Society."
%     \item Bijels Containing Magnetic Particles: A Simulation Study [Langmuir, 2010, 26, 11, 7928-7936]: 
%     "Reprinted (adapted) with permission from {COMPLETE REFERENCE CITATION}. Copyright {YEAR} 
%     American Chemical Society."
%     \item Numerical simulations of particulate suspensions via a discretized Boltzmann equation. Part 2. Numerical results
%     \item Aerobijels: Ultralight Carbon Monoliths from Cocontinuous Emulsions 
%     [Adv. Funct. Mater. 2020, 30, 1908383]: 5921080017718
%     \item Hierarchical assemblies of superparamagnetic colloids in 
%     time-varying magnetic fields [Soft Matter. 17, 5, 1120-1155]: 1552854-1
%     \item Colloidal Jamming at Interfaces: A Route to Fluid-Bicontinuous Gels [Science, 2005, 309, 5744]: 5966820525314
%     \item Capillary Interactions Between Anisotropic Colloidal Particles, [Phys. Rev. Lett. 2005, 94, 018301]: RNP/25/FEB/088185
% \end{itemize}