### 2025-03-21 chapters/paper3.tex

% Previous work into shearing bijels have demonstrated how particles prefer to move in the direction of shear and if strong enough, even detach from the interface. \cite{bonaccorso_shear_2020} It has also been shown how at curved interfaces, ellipsoidal particles

\section{Introduction}

While the microstructure and synthesis techniques for bijels have been explored in earnest, the rheology of bijels
has remained relatively unexplored. Some interesting phenomena for bijels and shearing include the 
presence of monogels, made by remixing the phase separated liquid domains of the bijel while the electrostatic
interactions between particles maintains the structure of the particle monolayer. \cite{sanz_colloidal_2009} 
Monogel formation has been observed in lutidine/water and styrene/butadiene oligomer based bijels but not ethane-diol nitromethane based bijels
\cite{sanz_colloidal_2009, bai_dynamics_2015, tavacoli_novel_2011} Computational Investigations have identified that confined bijels under shear undergo elongation of
the domains in the direction of shear succeeded by particle detachment from the interface and eventual failure of the bijel. \cite{bonaccorso_shear_2020}
More recently, bijels have been identified to be 2D glasses percolating in 3D space, characterized through comparing the complex rheology of bijels against
colloidal gels made from silica particles with electrostatic interactions. \cite{ching_bijel_2022}

In many of the manufacturing techniques outlined the rheological properties are essential in ensuring that 
the casting mixture remains processable. \cite{haase_continuous_2015,haase_situ_2016} In STrIPS, the flow rate of the casting mixture and the co-solvent
control the microstructure of the bijel and the morphology of the resulting bijel material, controlling whether droplets or ropes are synthesized. \cite{haase_continuous_2015}
When investigating the effect of magnetic fields on the rheological properties of suspensions, it has been shown that the viscosity of ferrofluids increases substantially due
to the self assembly of particles restricting the flow of fluid, greatly increasing the viscosity of the ferrofluid. cite{qiao_magnetorheological_2012} The application
of shear onto suspensions can also cause shear banding as interparticle interactions mediate particle fluid interactions, creating localized velocity differences to the bulk.
\cite{xu_relation_2013} Looking into emulsions, the application of magnetic fields increases the viscosity of emulsions, making them more shear thinning than when no field 
is applied.

In past chapters, we identified the microstructure and particle arrangement of bijels stabilized by ellipsoidal particles under magnetic fields and demonstrated
particles morphology specific ordering to the applied magnetic field. We also show that the local ordering of the particles is affected from the application of
the magnetic field. In this chapter, We probe the dynamics of bijels under constant shear to understand how these microstructural and particle monolayer changes
affect the rheology of bijels stabilized by ellipsoidal particles under magnetic fields. First, we define a shear capillary number, $Ca_s$
which in the literature lies 0.04 and 0.16. \cite{frijters_effects_2012, yang_capillary_2022}
However, these capillary number ranges would exceed the 
maximum mach number the model allows if this same range were used $(Ma \leq 0.03)$. To accommodate this limitation the 
largest capillary number used will be $Ca_s = 10^-4$ which correspond to a maximum of $Ma \approx 0.02$. As all particles used 
have only hard-sphere type interactions, it is expected that the behaviour seen should mimic 2D colloidal glasses 
percolating in 3D space, akin to what Ching and Mohraz saw, with an additional dependence on the direction of shear. 
This would predict the discovery of particle monolayer dependent elasticity and yield stress along with shear thinning 
behavior of the bijels.

To verify these predictions, bijel microstructure will be defined using four processing histories; The first is of a 
bijel simulated under a $\Bar{B} = 1$ magnetic field strength, the second is a bijel stimulated under no field, 
followed by the application of a $\Bar{B} = 1$ field after jamming, the third is a bijel simulated under no field, 
while the final microstructure is a bijel simulated under $\Bar{B} = 1$ magnetic field, followed by switching the 
field off after jamming. This gives insight into the impact of processing history on the shear properties of a bijel, 
in addition to the microstructural and colloidal insights gained. Based on the results in Bonaccorso et al., there 
should be shear driven and shear rate dependent elongation of the domains in the direction parallel to the applied 
shear which in this system will be seen as a reduction in $L_{\perp}$ and an increase in $L_{\parallel}$. 
\cite{bonaccorso_shear_2020} The microstructure anisotropy will also be a factor in the viscosity results, as 
larger domains are more permeable than smaller ones, meaning that bijels where $L_{\perp} > L_{\parallel}$ should 
see less of the domain elongation effects shown in Bonaccorso et al. as the permeability of the bijels rises with 
larger domain size. \cite{bonaccorso_shear_2020}

It is also expected that the effective viscosity will be different between the four microstructures dependent upon the 
degree of nematic ordering and microstructural anisotropy of the bijel. Nematic ordering of the particles in the direction of shear
should resulting in a lowered effective viscosity compared to bijels without nematically 
ordered particles. \cite{xu_relation_2013, vermant_flow-induced_2005} Tracking of the proportion of particles on the 
interface will also yield insight into how the packing of the particles affects the rate at which particles will get 
ejected from the interface. Bonnacorso et al. identified that shear applied onto bijels caused particles to align to the direction of
shear before being ejected from the interface. Systems with a larger $\eta_{eff}$ are predicted to have the largest domains, largest 
difference between initial and final particle order and lowest number of particles left on the interface once steady 
state has been established.

\section{Results}\label{sec:results_p3}

We begin this investigation by analyzing the rheological response of bijels stabilized by ellipsoidal particles under constant shear.
Specifically, we investigate the effect that particle order has on the rheological response characterized. We are also interested in
identifying the impact that the application of magnetic fields have on the rheological response of the bijel. 
Bijel templates simulated in Aim 1 with no field and a field strength of $\bar{B} = 1$ stabilized with a prolate and oblate particles with a 
particle volume fraction $\phi_p = 0.15$ were used as the starting point for these simulations. Next, we apply a field strength of
$\bar{B} = 1$ or switch off the applied field to all structures. This results in four processing setups; $\bar{B}:0\rightarrow 0$, $\bar{B}:0\rightarrow 1$,
$\bar{B}:1\rightarrow 0$ and $\bar{B}:1\rightarrow 1$. We then define a shear capillary number,  $Ca_s = \frac{\dot{\gamma} R_{s} \eta_{f}}{\sigma}$ where 
$\dot{\gamma}$ is the applied strain rate, $R_s$ is the equivalent spherical radius of the particles, $R_s = 7.9$, $\eta_f$ is the dynamic viscosity of the 
fluid and $\sigma$ is the surface tension of the interface. We use five shear capillary numbers, 
$Ca_s = 1\cdot10^{-4}, 5\cdot10^{-5}, 1\cdot10^{-5}, 5\cdot10^{-6}, 1\cdot10^{-6}$.

\subsection{Viscosity measurement}

The ordering of the particles play an important role in determining the rheology of suspensions as ordered particles can slip over one another
before the structure breaks. The adsorption of particles on the interface of a bijel could affect how this slip occurs. Previous work investigating
the rheology of bijels have determined shear thinning behavior characterized with a Herschel-Buckley fit equation. \cite{macmillan_rheological_2019} 
After applying shear on the bijel templates, we characterize and plot the time evolution of the the shear stress in Figure \ref{fig:stress_time}.

\begin{figure} 
    \centering 
    \includegraphics[scale=0.3]{../figures/results/paper3/stress-time_compare.png} 
    \caption{Time evolution of the shear stress of bijels stabilized with ellipsoidal particles as a function of the applied strain rate as
             a function of the initial microstructure and applied magnetic field. The observed stress response is difficult to say when a steady
             state is reached.} 
    \label{fig:stress_time} 
\end{figure}

Figure \ref{fig:stress_time} shows that there is no definitive steady state value of the shear stress obtained in the simulations performed. This is
attributed to the coarsening of the domains and reorientation of particles at the interface. \cite{tavacoli_novel_2011,macmillan_rheological_2019} 
Tavacoli et al. investigated the rheological response of Ethanediol/Nitromethane bijels and identified that upon placement of a needle to calculate the yield
stress of the bijel, the structure of the material was irreversibly altered. \cite{tavacoli_novel_2011} Macmillian et al. identified that repeated 
shearing of a bijel synthesized using direct mixing resulted in degraded mechanical performance. \cite{macmillan_rheological_2019}
To characterize the fit, we identify pseudo-steady state regions within each of the shear stress plots and fit the stress identified within this 
pseudo-steady state region to the strain rate. Pseudo-steady state regions are identified by measuring where the rate of change of shear stress is 
smallest for the longest duration. We fit the calculated stress to the applied strain rate using a Herschel-Buckley rheology model defined as
$\sigma_{xz} = \sigma_{y} + K(\dot{\gamma})^{n}$. $\sigma_{xz}$ is the shear stress calculated from the simulation, $\sigma_{y}$ is the yield stress of the bijel, 
$K$ is the flow consistency index and $n$ is the flow index. Where appropriate, we include results corresponding to a shear capillary number of 
$Ca_s = 10^{-6}$ as in many stress responses, $Ca_s = 10^{-6}$ may not always yield usable results.

\begin{figure} 
    \centering 
    \includegraphics[scale=0.4]{../figures/results/paper3/stress_strain-all.png} 
    \caption{Fitting the experimental shear stress to the strain rate using the Herschel Buckley model. We characterize the bijels are 
             shear thinning in all cases and that the bijels become less shear thinning as the particle order increases. We also
             demonstrate that the particle ordering affects the yield stress differently for particles stabilized by oblate and 
             prolate particles.} 
    \label{fig:stress_strain} 
\end{figure}

From Figure \ref{fig:stress_strain}, the slopes of the curve indicate a non-linear shear response to the applied strain rate, suggesting that the rheological properties
are non-Newtonian. Bijels stabilized by prolate and oblate particles also have different shear properties. For bijels stabilized by oblate particles, the stress curves
are more spread out suggesting that there is a larger variance in the properties arising from the change in the initial microstructure of the bijel. From the plot
the increase in particle order reduces the yield stress of the bijel. For bijels stabilized by rod like particles, the stress response of each processing 
condition is closer together, indicating a smaller variation in the response to the applied magnetic field compared to oblate particles. The shear response upon application
of the magnetic field also suggest that increased ordering of prolate particles makes the bijel more resistant to shear. We quantify the Herschel-Buckley fit parameters and 
tabulate the results in Table \ref{table:rheology_fit}.

\begin{table}[h!]
    \centering
    \renewcommand{\arraystretch}{1.5}  % Increases row height
    \begin{tabular}{||c c c c c||} 
     \hline
     Processing & $\alpha$ & $n$ & $K \frac{\Delta m (\Delta t)^{n-2}}{\Delta x} $ & $\sigma_{y} \frac{\Delta m}{\Delta x (\Delta t)^2}$ \\ [0.5ex] 
     \hline\hline
     $\bar{B}: 0 \rightarrow 0$ & 0.5 & 0.511 & 0.999 & $7.168 \cdot 10^{-5}$ \\ 
     \hline
     $\bar{B}: 0 \rightarrow 1$ & 0.5 & 0.579 & 0.944 & $4.008 \cdot 10^{-5}$ \\
     \hline
     $\bar{B}: 1 \rightarrow 0$ & 0.5 & 0.642 & 0.888 & $4.945 \cdot 10^{-5}$ \\
     \hline
     $\bar{B}: 1 \rightarrow 1$ & 0.5 & 0.637 & 0.914 & $1.715 \cdot 10^{-5}$ \\
     \hline
     $\bar{B}: 0 \rightarrow 0$ & 2 & 0.561 & 0.965 & $4.618 \cdot 10^{-5}$ \\
     \hline
     $\bar{B}: 0 \rightarrow 1$ & 2 & 0.609 & 0.944 & $4.101 \cdot 10^{-5}$ \\
     \hline
     $\bar{B}: 1 \rightarrow 0$ & 2 & 0.597 & 0.899 & $5.025 \cdot 10^{-5}$ \\
     \hline
     $\bar{B}: 1 \rightarrow 1$ & 2 & 0.606 & 0.885 & $6.435 \cdot 10^{-5}$ \\ [1ex] 
     \hline
    \end{tabular}
    \caption{Hershel Buckley fit parameters for different processing conditions applied to bijels stabilized by ellipsoidal particles.}
    \label{table:rheology_fit}
\end{table}

The results in Table \ref{table:rheology_fit} show that bijels remain shear thinning regardless of the processing history and particle morphology.
The particle morphology dependent differences in the shear response characterized in Figure \ref{fig:stress_strain} are quantified in Table \ref{table:rheology_fit}
where the yield stress differences predicted is clear to see. The bijel templates with an applied field of $\bar{B} = 1$ have larger flow indices
than bijel templates made with no applied field. The same trend exists here where the application of a magnetic field makes the bijel more Newtonian.
The yield stress of bijels stabilized with oblate particles decrease as the initial order of the particles and the applied field onto the bijel increase. 
The yield stress of bijels stabilized by prolate particles increases as the initial order of the particles and the applied field onto the bijel increase. In previous aims, we 
characterized that application of magnetic fields created orientational ordering of the bijels and that the local ordering of the bijels change. We  
showed that as magnetic fields were applied, oblate particles have lower local ordering on the interface while for prolate particles it increases.
Past investigations have identified that particle monolayers buckle more easily as local particle ordering decreases. \cite{prakash_buckling_2024} 
They identified that local ordering can enhance stress distribution among particles, making locally ordered particle monolayers better at withstanding shear.
We visualize the ordering of the particles at the interface to the z direction for both particle morphologies at $Ca_s = 10^{-6}$ for the 
case $\bar{B}: 1 \rightarrow 0$.

\begin{figure} 
    \centering 
    \includegraphics[scale=0.4]{../figures/results/paper3/tz_concat_startB-1_endB-0.png} 
    \caption{Visualization of the particle monolayer of bijels stabilized by prolate and oblate ellipsoidal particles with processing history $\bar{B}: 1 \rightarrow 0$
             undergoing an applied shear of $Ca_s = 10^{-6}$. Particles with a deeper purple shade have a larger angle to the z axis. The top row shows the initial 
             configuration of the particles while the bottom row shows the final configuration of the particles. } 
    \label{fig:particle_viz_tz} 
\end{figure}

In Figure \ref{fig:particle_viz_tz}, the particles are initially ordered to the direction of the applied magnetic field. Upon application of shear and removal
of the magnetic field, We see that the particles begin reorienting away from the z direction. We also observe that the reorientation of particles uncovers the interface.
The reorientation observed is particle morphology specific with prolate particles seeing more reorientation away from the applied shear direction than oblate particles.
Looking at the images, we see that prolate particles rotate away from the direction of the magnetic field and applied shear towards the direction of the shear gradient,
while oblate particles reorient towards minimizing area to the applied shear direction. While these snapshots show the initial and final orientations of the particles at the
interface, the time evolution of the particle monolayer changes to identify its role in the rheological properties characterized.

\subsection{Particle properties}

Particles at the interface of bijels stabilized with ellipsoidal particles reorient at the interface to minimize interfacial and
steric energies. \cite{gunther_timescales_2014} The particles stabilizing bijels in confined systems under shear have been shown to
orient to the direction of shear and ejected from the interface at sufficiently long times or shear.\cite{bonaccorso_shear_2020} 
We probe these aspects of the particle monolayer in this section. First, we plot the orientation of the particle to each cartesian axis 
to identify if shear induced ordering is present.

\begin{figure} 
    \centering 
    \includegraphics[scale=0.3]{../figures/results/paper3/angle_to_cartesian-SS.png} 
    \caption{Comparisons of the initial and final particle orientations of the particles to each cartesian direction under
             four processing histories for bijels stabilized with oblate and prolate ellipsoids at multiple shear rates. We show
             shear dependent ordering of the particle monolayer if the initial particle monolayer begins with particle order.} 
    \label{fig:particle_orientation_cartesian_shear} 
\end{figure}

From Figure \ref{fig:particle_orientation_cartesian_shear} we see that with bijel templates simulated under $\bar{B} = 0$, the
particles adopt a random orientation. Upon application of shear, we see that there is little directional ordering to the 
shear direction in the simulation time observed. Upon application of the magnetic field in the $\bar{B}:0 \rightarrow 1$ case, 
the particles order to the direction of the magnetic field. Here we see that the prolate particles order more readily to the applied
field than the oblate particles, seen as a larger clustering of particles at the $\theta_z = 0$ corner of the triangle. In the 
$\bar{B}:1 \rightarrow 0$ case particle morphology dependent partial relaxation of the particles away from the magnetic field direction 
is observed. Directionality of the particles to the field is still observed, suggesting that the ordering of the particles is not
immediately removed once the field is switched off. In the processing history of $\bar{B}:1 \rightarrow 1$, particles are locked
into ordering to the magnetic field direction. However we also see that some particles are reoriented away from the field direction
due to the applied shear. Prolate particles have a smaller distribution of particle orientations than oblate particles, suggesting that
the initial orientation and ordering of particles to the shear applied underpin the response observed. As the particles are adsorbed on the 
interface, the orientation of the particles are influenced by a balance between magnetic forces from the field and capillary forces 
from the interface. We can qualitatively calculate the time evolution of the capillary interactions using the interfacial angle.

\begin{figure} 
    \centering 
    \includegraphics[scale=0.3]{../figures/results/paper3/psi-time_compare.png} 
    \caption{Comparisons of the interfacial angle of bijels stabilized by ellipsoidal particles as a function of 
             the applied shear rate at four different processing histories. We show that upon application of shear
             particles tilt out of the interface reduces the interfacial area stabilized. $\bar{B}: 0 \to 1$ shows 
             large variations in the interface angle.} 
    \label{fig:interface_angle_shear} 
\end{figure}

From Figure \ref{fig:interface_angle_shear}, we demonstrate that there is a slow increase in the interface angle over time for all cases
except for $\bar{B}: 0 \to 1$. For the $\bar{B}: 0 \to 1$ case, this has been shown in Chapter \ref{chapter:aim2} to be due to particle
reorientation to the applied magnetic field. For all other cases, we see an average increase in $\langle \psi \rangle$ over time, showing
that the particles under shear reorient out of the interface. Particles reorienting out of the interface can be due to shear induced interfacial 
stress, modifying the curvature of the interface and thus the lowest energy particle orientation to the interface or shear induced increases in
interparticle forces, causing the particle to tilt out of the interface to minimize interparticle forces. The latter 
mechanism can be characterized through measurements of the number of particles on the interface as increases in interparticle forces result in ejection of
particles from the interface.

% We can characterize the impact of the interfacial angle on the 
% particle packing using the radial distribution function (rdf).

% \begin{figure} 
%     \centering 
%     \includegraphics[scale=0.3]{../figures/results/paper3/rdf-SS.png} 
%     \caption{Comparisons of the initial and final radial distribution function of the particles of bijels stabilized by ellipsoidal 
%     particles as a function of the applied shear rate at four different processing histories. I show that there are minor differences
%     in the particle ordering at the interface, suggesting that the particle tilting affects the coverage of particles at the interface.} 
%     \label{fig:rdf_shear} 
% \end{figure}

% From Figure \ref{fig:rdf_shear}, I show that the change in the packing of the particles on the interface is not correlated to the change 
% in the interfacial angle of the particles on the interface when changing the applied field strength in all cases. I then characterize the proportion 
% of particles on the interface to investigate how the proportion of particles on the interface changes and its impact on the interfacial area.

\begin{figure} 
    \centering 
    \includegraphics[scale=0.3]{../figures/results/paper3/n_int-time_compare.png} 
    \caption{Comparisons of the proportion of particles on the interface of bijels ($n_{int}$) stabilized by ellipsoidal particles as a function of 
             the applied shear rate at four different processing histories. We show that upon application of shear, many cases have a 
             reduction in the number of particles on the interface while in others it stays the same. 
            %  For cases where $n_{int}$ reduces, I can 
            %  attribute the domain size coarsening to this effect.
             } 
    \label{fig:particles_interface_prop_shear} 
\end{figure}

From the results in Figure \ref{fig:particles_interface_prop_shear}, the impact of the initial microstructure on the number of particles on the interface
differs for bijels stabilized by oblate and prolate ellipsoids. We split this analysis on processing history. While keeping the applied magnetic field
the same, little differences in the behavior of $n_{int}$ is observed, with small changes also being characterized for $\langle \psi \rangle$. 
Looking at the case of $\bar{B}:0 \rightarrow 0$, there is little change in the number of particles at the interface upon application of shear. This implies that the 
particles remain strongly adsorbed on the
interface. The $\langle \psi \rangle$ results also show this as this parameter increases slowly. This suggests that the particles remain firmly adsorbed
on the interface. Looking to the $\bar{B}:0 \rightarrow 1$ case, the number of particles at the interface drops while $\langle \psi \rangle$ increases. This
trend affects both particles which suggests that shear can pull particles out of the interface due to the decreased
interfacial adsorption energy the particle experiences when rotating out of the interface to orient to the applied field. The $\bar{B}:1 \rightarrow 1$ and 
$\bar{B}:1 \rightarrow 0$ have strong dependences on particle morphology. While an increase in $\langle \psi \rangle$ can be seen in both processing histories 
for both particle morphologies, a decrease in $n_{int}$ can be seen for prolate particles with processing history $\bar{B}:1 \rightarrow 0$ and for 
oblate particles with processing history $\bar{B}:1 \rightarrow 1$. We visualize $\langle \psi \rangle$ at the initial and final timestep for bijels stabilized with
oblate and prolate particles with processing $\bar{B}:1 \rightarrow 0$.

\begin{figure} 
    \centering 
    \includegraphics[scale=0.4]{../figures/results/paper3/psi_concat_startB-1_endB-0.png} 
    \caption{Visualization of the particle monolayer of bijels stabilized by prolate and oblate ellipsoidal particles with processing history $\bar{B}: 1 \rightarrow 0$
             undergoing an applied shear of $Ca_s = 10^{-6}$. Particles with a deeper red shade have a larger angle to the interface. The top and bottom row show the 
             configuration of the particles at the initial and final timestep respectively.} 
    \label{fig:particle_viz_psi} 
\end{figure}

Figure \ref{fig:particle_viz_psi} shows that the increase in $\langle \psi \rangle$ seen in Figure \ref{fig:interface_angle_shear} occurs due to different
reasons for both particle morphologies. For oblate particles, this is caused by particles orienting such that their symmetry axis is parallel to the interface
while prolate particles experienced interparticle forces induced by the applied shear, resulting in particles being ejected from the interface when the magnetic 
field is switched off. When keeping the magnetic field on, the oblate particles are forced to stay in sub-optimal particle orientations to the interface, resuling
in ejection of particles from the interface. The presence of the magnetic field reduces shear induced interparticle interactions owing to the ordering of
particles to the applied field, resulting in the mostly unchanged $n_{int}$ values. A increase in $\langle \psi \rangle$ and decrease in $n_{int}$ indicates domain
can indicate coarsening of the fluid domains as the interfacial area is reduced. In the next section, we investigate the the role of the interface properties and the 
impact of particle ejection and tilting of particles out of the interface on the microstructure of bijels.

\subsection{Microstructure change}

Bonaccorso et al. identified that the spherical particle stabilizers in their bijel simulations oriented to the direction of shear before being
ejected from the interface. \cite{bonaccorso_shear_2020} This then led to the domains of the bijel to coalesce and eventually destruction of the
bicontinuous microstructure. \cite{bonaccorso_shear_2020} In our bijels, we characterize particle morphology and initial microstructure specific 
properties of the particle monolayer upon application of shear. 
In this section, we probe the effect of applied shear to the microstructure of magnetically
responsive bijels stabilized by ellipsoidal particles. To investigate the effect of the particle angle and the interfacial coverage of the bijel on the
microstructural evolution of the interface, we plot the area averaged gaussian curvature over time. This parameter will allow insight into the evolution 
of the interface, namely the presence of domain coarsening, coalscence or break-off events.

\begin{figure} 
    \centering 
    \includegraphics[scale=0.3]{../figures/results/paper3/gaussian-time_compare.png} 
    \caption{Plots of the time evolution of the area averaged Gaussian curvature $\langle K \Sigma^{-2} \rangle$ for four 
             processing histories applied to bijels stabilized with oblate and prolate ellipsoids. We demonstrate that the 
             magnitude of the gaussian curvature reduces for all systems except for $\bar{B}: 0 \to 1$.} 
    \label{fig:gaussian_curvature_time_shear} 
\end{figure}

Figure \ref{fig:gaussian_curvature_time_shear} reveals that the area averaged Gaussian curvature generally decreases in magnitude over time for all cases except 
$\bar{B}: 0 \to 1$ indicating that the increase in $\langle \psi \rangle$ causes domain coarsening of the bijel. $\bar{B}: 0 \to 1$ is
unique because particle rearrangements at the interface dominates the dynamics of the microstructure, not observed in all other
cases. From the time evolution of the curvature, the same mechanism drives coarsening in all cases. From the results in Figures \ref{fig:particles_interface_prop_shear} 
and \ref{fig:interface_angle_shear}, domain coarsening is driven through a reduction in the interfacial area covered by the particle monolayer. 
From the curvature results characterized in Figure \ref{fig:gaussian_curvature_time_shear}, the area averaged curvature coarsening is shear independent. This is further
evidence that the mechanism of coarsening is identical in each case.

\begin{figure} 
    \centering 
    \includegraphics[scale=0.3]{../figures/results/paper3/anisotropy_compare.png} 
    \caption{Comparisons of the time evolution of the directional domain sizes that increase upon application of a magnetic field for bijels stabilized 
             by ellipsoidal particles undergoing various shear rates under four processes. The directional domain size increases in all cases except for when the
             field is switched off.} 
    \label{fig:domain_size_aniso_time_shear} 
\end{figure}

From Figure \ref{fig:domain_size_aniso_time_shear}, the directional domain sizes in the orientations with the maximum change were characterized. For oblate and prolate 
particles, this was found to be $L_{\perp}$ and $L_{\parallel}$ respectively. In all cases except for $\bar{B}: 1 \to 0$, this length scale increases matching the expected
trend that increasing $\langle \psi \rangle$ matches domain coarsening. The decrease in these length scales for $\bar{B}: 1 \to 0$ suggest that the shear induced
reordering of interfacial particles after application of a field affects the evolution of the microstructure. For both particles, this is due to particle mediated
shear forces causing domains to coarsen more in specific directions. From the snapshots of the particle arrangement in Figure \ref{fig:particle_viz_psi} the particles were
shown to reorient at the interface. This process did not occur for $\bar{B}: 1 \to 1$ as the presence of magnetic fields acted to keep particles aligned a specific
direction.

% Previous investigations have 
% demonstrated how the orientation and arrangement of particles on the interface affect the domain size. 
% \cite{gunther_timescales_2014} I focus our investigation on this aspect in the following plot.

% \begin{figure} 
%     \centering 
%     \includegraphics[scale=0.3]{../figures/results/paper3/domain_size-time_compare.png} 
%     \caption{Comparisons of the time evolution of the average domain size for bijels stabilized by ellipsoidal particles undergoing
%              various shear rates under four processes. Initially, the domain size coarsens for all runs under applied shear before
%              reducing in some cases.} 
%     \label{fig:domain_size_time_shear} 
% \end{figure}

% From Figure \ref{fig:domain_size_time_shear}, domains coarsen with time in all cases. When increasing the applied magnetic field in the case of 
% $\bar{B}:0 \rightarrow 1$, I see shear assisted domain coarsening characterized as a 20 \% and 10 \% increase in the domain size of bijels
% stabilized by oblate and prolate particles respectively. This is approximately 5 times greater than the response observed in Figure \ref{fig:domain_size-field_on}
% In the  $\bar{B}: 1 \to 0$ and $\bar{B}: 1 \to 1$ cases, an initial plateau in domain size is observed, followed by a divergence in 
% growth depending on $Ca$. The shear rate appears to influence domain coarsening, with higher $Ca_s$ leading to delayed or 
% suppressed growth, especially in the presence of a field. The difference between applying $\bar{B}: 0 \to 1$ and 
% removing $\bar{B}: 1 \to 0$ the field demonstrates the stability of the particle monolayer when a field is switched off.
% There is little difference in the qualitative trends of domain size change for bijels stabilized with oblate or prolate particles.
% To characterize whether the domain size coarsening observed arises from different phenomena in each case, the area averaged curvature is plotted.

\section{Conclusions}

% In this chapter, I investigated the rheology of bijels stabilized by ellipsoidal particles under shear, focusing on the influence of initial microstructure and the 
% application of a constant magnetic field on their shear response. The shear behavior was characterized using a Herschel-Bulkley fitting function to extract the flow index, 
% consistency index, and yield stress. My analysis confirmed that the shear-thinning nature of bijels remains robust, regardless of particle morphology or applied magnetic 
% field. However, the flow index and yield stress trends were strongly dependent on these factors.  
% I found that increased particle ordering led to distinct yield stress behaviors: oblate particles resulted in a lower yield stress, while prolate particles exhibited a 
% higher yield stress. When examining the properties of the particle monolayer, I identified a clear impact of particle morphology on interparticle interactions at the 
% interface. In the absence of a magnetic field, both particle monolayers remained adsorbed at the interface. However, when a magnetic field was initially applied, the 
% behavior diverged significantly oblate particles flipped out of the interface to minimize interfacial energy, whereas prolate particles tilted out. These differences 
% directly influenced the shear response, altering how the surrounding fluid flows around the particles and leading to the observed non-Newtonian behavior and yield 
% stress characteristics.  
% Furthermore, I investigated the microstructural effects of shear and magnetic field-induced particle ordering and found that coarsening was the primary driver of 
% domain size evolution in sheared bijels. This coarsening was facilitated by a reduction in interfacial particle coverage. When analyzing the anisotropic domain size evolution, 
% I observed that in most cases, the direction of greatest structural change aligned with expected trends—except when the magnetic field was switched off. This deviation is 
% attributed to differences in particle monolayer evolution, depending on whether the bijel was formed with or without an initially applied field.

The rheological properties of bijels are of importance in many fabrication schemes such as STrIPS and homogenization. \cite{haase_continuous_2015,cai_bijels_2017}
To characterize how magnetically responsive bijels can be used in these synthesis techniques, we studied the effect of particle morphology, initial microstructure and
applied magnetic fields on the rheology of bijels experiencing a constant shear. Bijel templates made with $\bar{B} = 0$ and $\bar{B} = 1$ were selected, with field strengths
of $\bar{B} = 0$ and $\bar{B} = 1$ applied to each template. This results in four processing histories per particle morphology namely $\bar{B}:0 \rightarrow 0$,
$\bar{B}:0 \rightarrow 1$, $\bar{B}:1 \rightarrow 0$ and $\bar{B}:1 \rightarrow 1$. We then apply a constant shear rate onto each bijel.

We first characterized the shear stress imparted by bijels under several strain rates. The shear stress calculated continued to increase in line with expectations
for colloidal glasses as ageing occurs. Thus a pseudo-steady state was selected to compare the shear properties of the bijel.
We fit the shear stress data using a Herschel Buckley fit and identify that the flow index of bijels stabilized by ellipsoidal 
particles under magnetic fields became less shear thinning. The yield stress of bijels stabilized by oblate particles was shown to reduce by 4x while the yield stress
of bijels stabilized by prolate particles was shown to increase by 2x upon application of magnetic fields.

To characterize these differences, closer attention is paid to the particle monolayer of the bijel. Shear induced reorientation of the ellipsoids can be observed in cases
where a magnetic field was used observed, characterized as the orientation of particles drift away from the magnetic field direction. When analyzing the effect this reorientation
has on the particles at the interface, the interface angle in all cases increases as a function of the shear rate. However the quantitative behavior of oblate particles differ
from prolate particles. Under magnetic fields, oblate particles tend to flip out of the interface while prolate particles tilt out of the interface. The impact that
the particle tilting and flipping out of the interface can also be seen in the number of particles in the interface, where the differences in the particle morphology 
can be characterized as differences in the number of particles at the interface in response to the magnetic field.

A reduction in the number of particles at the interface and the particles tilting out of the interface reduces the interfacial coverage. The reduction in interfacial
coverage manifests as domain coarsening, characterized as a smooth decrease in the magnitude of the area averaged Gaussian curvature. The changes in the particle monolayer
properties manifest as changes in the evolution of the directional domain size evolution due to shear induced particle reorientation at the interface.

Past literature
has shown how bijels are susceptible to failure when being used as crossflow reactors due to shear induced domain coarsening resulting in microstructure failure. 
\cite{boakye-ansah_controlling_2020} The investigations conducted here demonstrate that controlling the orientations of particles at interfaces are crucial to 
the rheological properties of bijels. Other properties not considered here such as different particle aspect ratios, the effect of surface tension on the shear response of
bijels and the volume fraction of particles are not considered here. Additionally, complex rheology of the bijels can determine detailed rheological properties such as the
gel point and phase difference of the bijel which can yield further rheological characterization of bijels stabilized by magnetically responsive ellipsoidal particles.