### 2025-04-08 licenses/licenses.tex

\section{License numbers}

\begin{itemize}
    \item E. Garcia, C. S. Wang, R. N. Sanderson, K. M. McDevitt, Y. Zhang, L. Valdevit, D. R.
    Mumm, A. Mohraz and R. Ragan. Scalable synthesis of gyroid-inspired freestanding three-
    dimensional graphene architectures. 1, 3870. Reproduced using license number 1548092-1
    \item Stratford, R. Adhikari, I. Pagonabarraga, J.-C. Desplat and M. E. Cates. Colloidal jam-
    ming at interfaces: A route to fluid-bicontinuous gels. 309, 2198. Reproduced using license number 5966820525314
    \item F. Haase, K. J. Stebe and D. Lee. Continuous fabrication of hierarchical and asymmetric
    bijel microparticles, fibers, and membranes by solvent transfer-induced phase separation
    (STRIPS). 27, 7065. Reproduced using license number 5913140219015
    \item K. Tham, W. M. Ng, S. S. Leong, S. P. Yeap, S. C. Low, H. L. Lee and J. Lim. Mag-
    netophoresis of magnetic pickering emulsions under low field gradient: Macroscopic and
    microscopic motion. 37, 1811: "Reprinted (adapted) with permission from 
    Langmuir 2021, 37, 5, 1811-1822. Copyright 2025 American Chemical Society."
    \item G. B. Davies, T. Kr¨uger, P. V. Coveney, J. Harting and F. Bresme. Assembling ellipsoidal
    particles at fluid interfaces using switchable dipolar capillary interactions. 26, 6715. Reproduced under the Creative Commons CC BY License.
    \item C. Loudet, A. M. Alsayed, J. Zhang and A. G. Yodh. Capillary interactions between
    anisotropic colloidal particles. 94, 018301. Reproduced using license number RNP/25/FEB/088185
    \item S. S. Velankar. A non-equilibrium state diagram for liquid/fluid/particle mixtures. 11, 8393. Reproduced using license number 1551968-1
    \item Bonaccorso, S. Succi, M. Lauricella, A. Montessori, A. Tiribocchi and K. H. Luo. Shear
    dynamics of confined bijels. 10, 095304.: Reproduced under the Creative Commons CC BY License.
    \item Particle-Stabilized Emulsions and Colloids: Formation and Applications; Ngai, T., Bon, S. A. F., Eds.; RSC soft matter series; 
    Royal Society of Chemistry, RSC Publ: Cambridge, 2015: Reproduced under the Royal Society of Chemistry license number 1597404-1
    \item Particle schematic at an interface demonstrating the various quantities of interest. Reprinted Figure 1 from
    Guillen-Chable, F.; Bayona, A.; Rodríguez-Zapata, L. C.; Castano, E. Phase Separation of Intrinsically Disordered 
    Nucleolar Proteins Relate to Localization and Function. IJMS 2021, 22 (23), 13095. with permission under the Creative Common CC BY license.
    \item Cavallaro, M.; Botto, L.; Lewandowski, E. P.; Wang, M.; Stebe, K. J. Curvature-Driven Capillary Migration and Assembly of Rod-like Particles. 
    Proc. Natl. Acad. Sci. U.S.A. 2011, 108 (52), 20923-20928. Reproduced under the CC-BY license
    \item Hijnen, N.; Cai, D.; Clegg, P. S. Bijels Stabilized Using Rod-like Particles. Soft Matter 2015, 11 (22), 4351-4355. Reproduced under the CC-BY license
    \item Wagner, N. J.; Brady, J. F. Shear Thickening in Colloidal Dispersions. Physics Today 2009, 62 (10), 27-32. Reproduced under the CC-BY license
    \item Cooke, M. E.; Rosenzweig, D. H. The Rheology of Direct and Suspended Extrusion Bioprinting. APL Bioengineering 2021, 5 (1), 
          011502. Reproduced under the CC BY license.
    \item S. Schmieschek, L. Shamardin, S. Frijters, T. Kr¨uger, U. D. Schiller, J. Harting and P. V.
    Coveney. LB3D: A parallel implementation of the Lattice-Boltzmann method for simulation
    of interacting amphiphilic fluids. Computer Physics Communications 217, 149 (2017). Reproduced under the Creative Commons CC BY License
\end{itemize}

% \begin{itemize}
%     \item Continuous Fabrication of Hierarchical and Asymmetric Bijel Microparticles, 
%     Fibers, and Membranes by Solvent Transfer‐Induced Phase Separation (STRIPS): 5913140219015
%     \item Scalable synthesis of gyroid-inspired freestanding three-dimensional graphene architectures: 1548092-1
%     \item State diagram for particle stabilized emulsions. [Soft Matter, 2015,11, 8393-8403]: 1551968-1
%     \item Scalable Manufacturing of Hierarchical Biphasic Bicontinuous Structures via Vaporization-Induced 
%     Phase Separation (VIPS) [Materials Letters 2020, 2, 5, 524-530]: "Reprinted (adapted) with permission from 
%     {COMPLETE REFERENCE CITATION}. Copyright {YEAR} American Chemical Society."
%     \item Modulation of Cellulose Nanocrystals Amphiphilic Properties to Stabilize Oil/Water Interface 
%     [Biomacromolecules 2016, 17, 5, 1748–1756]: "Reprinted (adapted) with permission from 
%     {COMPLETE REFERENCE CITATION}. Copyright {YEAR} American Chemical Society."
%     \item Dextran-Based Nanoparticles to Formulate pH-Responsive Pickering Emulsions: A Fully Degradable 
%     Vector at a Day Scale [Biomacromolecules 2020, 21, 12, 5358–5368]: "Reprinted (adapted) with permission from 
%     {COMPLETE REFERENCE CITATION}. Copyright {YEAR} American Chemical Society."
%     \item Stable emulsions with thermally responsive microstructure and rheology using poly(ethylene oxide) star 
%     polymers as emulsifiers [Journal of Colloid and Interface Science, 2013, 394, 284-292]: 5921030909190
%     \item Magnetophoresis of Magnetic Pickering Emulsions Under Low Field Gradient: Macroscopic and Microscopic 
%     Motion [Langmuir 2021, 37, 5, 1811–1822]: "Reprinted (adapted) with permission from 
%     {COMPLETE REFERENCE CITATION}. Copyright {YEAR} American Chemical Society."
%     \item Influence of magnetic field on the orientation of anisotropic magnetic particles at liquid interfaces 
%     [Phys Chem Chem Phys 2014, 16, 47, 26051-26058]: 1551970-1
%     \item Pickering Emulsions with Controllable Stability [Langmuir, 2005, 21, 6, 2158-2162]: "Reprinted (adapted) with permission from 
%     {COMPLETE REFERENCE CITATION}. Copyright {YEAR} American Chemical Society."
%     \item Bijels Containing Magnetic Particles: A Simulation Study [Langmuir, 2010, 26, 11, 7928-7936]: 
%     "Reprinted (adapted) with permission from {COMPLETE REFERENCE CITATION}. Copyright {YEAR} 
%     American Chemical Society."
%     \item Numerical simulations of particulate suspensions via a discretized Boltzmann equation. Part 2. Numerical results
%     \item Aerobijels: Ultralight Carbon Monoliths from Cocontinuous Emulsions 
%     [Adv. Funct. Mater. 2020, 30, 1908383]: 5921080017718
%     \item Hierarchical assemblies of superparamagnetic colloids in 
%     time-varying magnetic fields [Soft Matter. 17, 5, 1120-1155]: 1552854-1
%     \item Colloidal Jamming at Interfaces: A Route to Fluid-Bicontinuous Gels [Science, 2005, 309, 5744]: 5966820525314
%     \item Capillary Interactions Between Anisotropic Colloidal Particles, [Phys. Rev. Lett. 2005, 94, 018301]: RNP/25/FEB/088185
% \end{itemize}